---
title: "Initial Korean covid 19"
author: "Jasmin Quintanilla"
date: "Summer 2024"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```

# Introduction

Our file is KRdata.csv  
dataset from South Korea 1 Jan to 1 May 2020  
Is dying independent of sex?  

### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female  
77 of the data entries were coded sex  as blank, so we omit these.  
state of a patient must be died or alive - we adjust that as well     
  
# Your Question - Is dying independent of sex?

$H_0$:There is no difference in the death rates deeding on on sex

$H_A$: There is a difference in the death rates are higher for males 

# Methods

Result is a categorical variable, sex is categorical.   

Result depends on sex, we use helper file cat~cat  

#Results


## Graphical results - use ggplot commands to do this 

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

This bar chart shows that there's more death in males than females. 

## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

This numerical summary shows that 64% of male died more than females which is 36%.


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The p-value for the Chi-squared is 0.001, so we reject the null hypothesis. 

The fisher exact test shows that men are 2.3 times more likely to die than women in Korea.


## Conclusions

The original question was if there was a relationship between sex and deaths in koren due to covid.


